Forage site distributions quantify collective foraging in honey bee colonies: Methods

### 2022-04-11 analysis/fit_models_durations.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

alldata <- data %>%
  select(date, site, duration.seconds) %>%
  rename(foraging_distance = duration.seconds)
```

@@

# set up bounds for the collective model
p_bnds <- c(0, 1.0)
bs_bnds <- c(1.0e-6, 10)
br_bnds <- c(1.0e-6, 10)
as_bnds <- c(1.0e-12, 10)
ar_bnds <- c(1.0e-12, 10)
collective_bounds <- rbind(
  p_bnds, bs_bnds,
  br_bnds, as_bnds,
  ar_bnds
)

@@

# save results
saveRDS(df, file = "results/site_fit_results.Rda")

>>>>

alldata <- data %>%
  filter(Year == 2017) %>% # remove pilot data conducted in 2016
  select(date, site, duration.seconds) %>%
  rename(foraging_distance = duration.seconds)
```

@@

@@

# save results
# saveRDS(df, file = "results/site_fit_results.Rda")

### 2022-04-11 analysis/landuse_effect_jackknife.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

# put proportion of scouts into landuse data
urban_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu)
urban_lu_long <- urban_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu)
rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

@@

plt <- ggplot(data = urban_jk_loadings, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_point(aes(colour = landuse_type)) +
  geom_segment(aes(
    x = landuse_type, xend = landuse_type,
    y = 0, yend = loadings, colour = landuse_type
  )) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -urban_cuttoff,
    ymax = urban_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text(size = 12),
    strip.text.x = element_text(size = 18),
    axis.title.x = element_text(size = 18)
  ) +
  scale_fill_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  ) +
  facet_wrap(~site_removed, ncol = 2)
ggsave(
  plot = plt,
  filename = "results/figures/SI_figures/urban_jk_individual_loadings.png",
  width = 120,
  height = 200,
  units = "mm",
  dpi = 300
)
plt
```

>>>>

# put proportion of scouts into landuse data
urban_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(urban_lu)

urban_lu_long <- urban_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu) %>%
  filter(site != "ROT") # remove ROT as fit is bad

rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")

```

@@

plt <- ggplot(data = urban_jk_loadings, aes(
  x = landuse_type, y = loadings, fill = landuse_type
)) +
  geom_point(aes(colour = landuse_type)) +
  geom_segment(aes(
    x = landuse_type, xend = landuse_type,
    y = 0, yend = loadings, colour = landuse_type
  )) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5)
  ) +
  annotate(
    "rect",
    ymin = -urban_cuttoff,
    ymax = urban_cuttoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text(size = 15),
    strip.text.x = element_text(size = 18),
    axis.title.x = element_text(size = 18)
  ) +
  scale_fill_manual(
    values = c(
      "arable" = "#CCBB44", "built.up.area" = "#EE6677",
      "Non.agricultural.unmanaged.green.space" = "#228833",
      "water" = "#4477AA", other = "#BBBBBB"
    )
  ) +
  facet_wrap(~site_removed, ncol = 2)
ggsave(
  plot = plt,
  filename = "results/figures/SI_figures/urban_jk_individual_loadings.png",
  width = 120,
  height = 200,
  units = "mm",
  dpi = 300
)
plt
```

### 2022-04-11 analysis/landuse_effect_prop_scouts.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu)
rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

@@

# make a scree plot
ggplot(
  data.frame(PC = seq(1, 10), varpercent = varpercent),
  aes(x = PC, y = varpercent)
) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 1 / ncol(urban.pcadata) * 100, col = "red") +
  ylab("% Variance") +
  scale_x_continuous(breaks = seq(1, 10))
```

@@

# make fit map with agrirural landuses
agrirural_plotdata$site <- rural_lu$site
agrirural_plotdata$prop_scouts <- rural_lu$p
ar_sortedpropscouts <- arrange(agrirural_plotdata, prop_scouts)

@@

# select sites to draw attention to, those sitting on the trend
ar_marked <- ar_sortedpropscouts %>%
  filter(site %in% c("MEL", "ROT", "BEL"))

@@

# make the maps for each of the sites selected in the previous step
site_shape_data_ar <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/agri-rural/",
    ar_marked[i, "site"], "/", ar_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  # add major LU categories
  df_merged <- df %>%
    mutate(descriptio = toupper(descriptio)) %>%
    full_join(agrirural_luconversions, by = c("descriptio" = "code"))
  site_shape_data_ar[[ar_marked[i, "site"]]] <- df_merged
  site_plots[[ar_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df_merged,
      aes(fill = category)
    ) +
    theme_void() +
    scale_fill_manual(
      values = c(
        "arable" = "#CCBB44", "built-up area" = "#EE6677",
        "Non-agricultural unmanaged green space" = "#228833",
        "water" = "#4477AA", "pasture" = "#BBBBBB", "woodland" = "#BBBBBB",
        "Non-agricultural managed green space" = "#BBBBBB", "fruit" = "#BBBBBB",
        "OSR" = "#BBBBBB", "other agricultural" = "#BBBBBB"
      )
    )
}

@@

loading_plot <- ggplot(
  data = argi.rural.loadings,
  aes(
    x = Landuse, y = Correlation
  )
) +
  geom_boxplot(aes(fill = Landuse), lwd = 0.1) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = rural_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  annotate(
    "rect",
    ymin = -cutoff,
    ymax = cutoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Arable" = "#CCBB44", "Built Up Area" = "#EE6677",
      "NAUMGS" = "#228833",
      "Water" = "#4477AA", Other = "#BBBBBB"
    )
  )

@@

# make the maps for each of the sites selected in the previous step
site_shape_data_u <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/urban/",
    u_marked[i, "site"], "/", u_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  # add major LU categories
  df_merged <- df %>%
    mutate(descriptio = toupper(descriptio)) %>%
    full_join(urban_luconversions, by = c("descriptio" = "code"))
  site_shape_data_u[[u_marked[i, "site"]]] <- df_merged
  site_plots[[u_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df_merged,
      aes(fill = category)
    ) +
    theme_void() +
    scale_fill_manual(
      values = c(
        "continuous central" = "#BBBBBB", "dense residential" = "#BBBBBB",
        "sparse residential" = "#EE6677", "amenity grassland" = "#BBBBBB",
        "water" = "#BBBBBB", "parks allotments cemeteries" = "#228833",
        "Railway" = "#CCBB44", "woodland" = "#BBBBBB"
      )
    )
}

@@

loading_plot <- ggplot(
  data = urban.loadings,
  aes(
    x = Landuse, y = Correlation
  ),
  fill = Landuse
) +
  geom_boxplot(aes(fill = Landuse), lwd = 0.1) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  annotate(
    "rect",
    ymin = -cutoff,
    ymax = cutoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Continuous Central" = "#BBBBBB", "Dense Residential" = "#BBBBBB",
      "Sparse Residential" = "#EE6677", "Railway" = "#CCBB44",
      "Parks Allotments Cemeteries" = "#228833",
      "Water" = "#BBBBBB", Other = "#BBBBBB"
    )
  )

@@

joined <- bind_rows(rural_lu, urban_lu)
joined[is.na(joined)] <- 0

joined.pcdata <- joined[, 3:length(joined)]

n <- length(joined$p)
y <- (joined$p * (n - 1) + 0.5) / n
joined.pcdata$p_transform <- y

model <- plsRbeta(
  p_transform ~ .,
  data = joined.pcdata,
  nt = ncol(joined.pcdata) - 1,
  model = "pls-beta",
  verbose = FALSE
)

print(model)

plsbetamodel <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4,
  plsdata
)

newdat <- data_frame(
  Comp_.1 = seq(-50, 140, 0.1),
  Comp_.2 = seq(-50, 140, 0.1),
  Comp_.3 = seq(-50, 140, 0.1),
  Comp_.4 = seq(-50, 140, 0.1)
)

plotdata <- plsdata
plotdata$p <- joined$p

plotdata$LU <- ""
plotdata$LU[1:10] <- "Rural"
plotdata$LU[11:20] <- "Urban"

ggplot(
  data = plotdata,
  aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)
) +
  geom_point(aes(colour = LU)) +
  geom_rug(aes(colour = LU)) +
  scale_color_manual(values = c("Rural" = "black", "Urban" = "red")) +
  geom_line(data = newdat, aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = fit)) +
  xlim(min(-5.5), max(5.5))

ggplot(
  data = plotdata,
  aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)
) +
  geom_boxplot(aes(group = LU))
```

>>>>

rural_lu <- sitedata %>%
  select(site, p) %>%
  inner_join(rural_lu)  %>%
  filter(site != "ROT") # remove ROT as fit is bad

rural_lu_long <- rural_lu %>%
  pivot_longer(!c(site, p), names_to = "landuse", values_to = "values")
```

@@

# make a scree plot
ggplot(
  data.frame(PC = seq(1, 9), varpercent = varpercent),
  aes(x = PC, y = varpercent)
) +
  geom_bar(stat = "identity") +
  geom_hline(yintercept = 1 / ncol(urban.pcadata) * 100, col = "red") +
  ylab("% Variance") +
  scale_x_continuous(breaks = seq(1, 10))
```

@@

# make fit map with agrirural landuses
agrirural_plotdata$site <- rural_lu$site
agrirural_plotdata$prop_scouts <- rural_lu$p
ar_sortedpropscouts <- arrange(agrirural_plotdata, prop_scouts) %>%
  filter(site != "ROT")

@@

# select sites to draw attention to, those sitting on the trend
ar_marked <- ar_sortedpropscouts %>%
  filter(site %in% c("MEL", "SRA", "HER"))

@@

# make the maps for each of the sites selected in the previous step
site_shape_data_ar <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/agri-rural/",
    ar_marked[i, "site"], "/", ar_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  # add major LU categories
  df_merged <- df %>%
    mutate(descriptio = toupper(descriptio)) %>%
    full_join(agrirural_luconversions, by = c("descriptio" = "code"))
  site_shape_data_ar[[ar_marked[i, "site"]]] <- df_merged
  site_plots[[ar_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df_merged,
      aes(fill = category)
    ) +
    theme_void() +
    scale_fill_manual(
      values = c(
        "arable" = "#018B27", "built-up area" = "#ED6778",
        "Non-agricultural unmanaged green space" = "#BEC330",
        "water" = "#536CAB", "pasture" = "#BBBBBB", "woodland" = "#BBBBBB",
        "Non-agricultural managed green space" = "#BBBBBB", "fruit" = "#BBBBBB",
        "OSR" = "#BBBBBB", "other agricultural" = "#BBBBBB"
      )
    )
}

@@

loading_plot <- ggplot(
  data = argi.rural.loadings,
  aes(
    x = Landuse, y = Correlation
  )
) +
  geom_boxplot(aes(fill = Landuse), lwd = 0.1) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = rural_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  annotate(
    "rect",
    ymin = -cutoff,
    ymax = cutoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Arable" = "#018B27", "Built Up Area" = "#ED6778",
      "NAUMGS" = "#BEC330",
      "Water" = "#536CAB", Other = "#BBBBBB"
    )
  )

@@

# make the maps for each of the sites selected in the previous step
site_shape_data_u <- list()
site_plots <- list()
for (i in seq_len(3)) {
  path <- paste0(
    "shapefiles/urban/",
    u_marked[i, "site"], "/", u_marked[i, "site"],
    ".shp"
  )
  df <- st_read(path) %>%
    st_transform(
      CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0")
    )
  # add major LU categories
  df_merged <- df %>%
    mutate(descriptio = toupper(descriptio)) %>%
    full_join(urban_luconversions, by = c("descriptio" = "code"))
  site_shape_data_u[[u_marked[i, "site"]]] <- df_merged
  site_plots[[u_marked[i, "site"]]] <- ggplot() +
    geom_sf(
      data = df_merged,
      aes(fill = category)
    ) +
    theme_void() +
    scale_fill_manual(
      values = c(
        "continuous central" = "#BBBBBB", "dense residential" = "#BBBBBB",
        "sparse residential" = "#ED6778", "amenity grassland" = "#BBBBBB",
        "water" = "#BBBBBB", "parks allotments cemeteries" = "#018B27",
        "Railway" = "#BEC330", "woodland" = "#BBBBBB"
      )
    )
}

@@

loading_plot <- ggplot(
  data = urban.loadings,
  aes(
    x = Landuse, y = Correlation
  ),
  fill = Landuse
) +
  geom_boxplot(aes(fill = Landuse), lwd = 0.1) +
  geom_hline(
    yintercept = 0, color = "black"
  ) +
  geom_text(
    data = urban_loading_text,
    aes(x = landuse_type, y = result, label = landuse_type),
    hjust = 0.5,
    vjust = 0.5,
    nudge_y = c(-0., -0.), nudge_x = c(0.5, 0.5), size = rel(textsize)
  ) +
  annotate(
    "rect",
    ymin = -cutoff,
    ymax = cutoff,
    xmin = 0, xmax = Inf,
    alpha = 0.2
  ) +
  coord_flip() +
  theme_classic() +
  theme(
    axis.line.y = element_blank(),
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    legend.position = "None",
    text = element_text()
  ) +
  scale_fill_manual(
    values = c(
      "Continuous Central" = "#BBBBBB", "Dense Residential" = "#BBBBBB",
      "Sparse Residential" = "#ED6778", "Railway" = "#BEC330",
      "Parks Allotments Cemeteries" = "#018B27",
      "Water" = "#BBBBBB", Other = "#BBBBBB"
    )
  )

@@

# joined <- bind_rows(rural_lu, urban_lu)
# joined[is.na(joined)] <- 0

# joined.pcdata <- joined[, 3:length(joined)]

# n <- length(joined$p)
# y <- (joined$p * (n - 1) + 0.5) / n
# joined.pcdata$p_transform <- y

# model <- plsRbeta(
#   p_transform ~ .,
#   data = joined.pcdata,
#   nt = ncol(joined.pcdata) - 1,
#   model = "pls-beta",
#   verbose = FALSE
# )

# print(model)

# plsdata <- data.frame(
#   prop.scouts = model$dataY,
#   model$tt
# )

# plsbetamodel <- betareg(
#   prop.scouts ~ Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4,
#   plsdata
# )

# summary(plsbetamodel)

# newdat <- data_frame(
#   Comp_.1 = seq(-50, 140, 0.1),
#   Comp_.2 = seq(-50, 140, 0.1),
#   Comp_.3 = seq(-50, 140, 0.1),
#   Comp_.4 = seq(-50, 140, 0.1)
# )

# newdat$fit <- predict(plsbetamodel, newdat, type = "response")

# plotdata <- plsdata
# plotdata$p <- joined$p

# plotdata$LU <- ""
# plotdata$LU[1:10] <- "Rural"
# plotdata$LU[11:20] <- "Urban"

# ggplot(
#   data = plotdata,
#   aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = p)
# ) +
#   geom_point(aes(colour = LU)) +
#   geom_rug(aes(colour = LU)) +
#   scale_color_manual(values = c("Rural" = "black", "Urban" = "red")) +
#   geom_line(data = newdat, aes(x = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4, y = fit)) +
#   xlim(min(-5.5), max(5.5))

# ggplot(
#   data = plotdata,
#   aes(x = LU, y = Comp_.1 + Comp_.2 + Comp_.3 + Comp_.4)
# ) +
#   geom_boxplot(aes(group = LU))
```

### 2022-04-11 analysis/manuscript/methods.Rmd
updated to use 2017 data, updated ms and si and methods. Lots of stuff as we approach final edits for submission

A circular environment is first created with radius $r = 2.5$. The number of flowers in the environment is generated as a random Poisson variable with rate equal to 5000 multiplied by the area of the environment. These are placed on polar coordinates with a uniformly selected angle, $\theta$, between 0 and 2$\pi$ and a radial value, $\rho$, between 0 and $r$, determined from the square root of the uniform position values multiplied by $r$. These polar coordinates are converted to Cartesian coordinates using the formulas $x = \rho\cos(\theta)$ and $y = \rho\sin(\theta$). Each location is then assigned to an instance of a flower object along with a random quality of between 0 and 10. This quality is combined with the distance of the flower to the centrally located hive to form a measure of how profitable the flower is (see model, equation?).

@@

The simulation was run 100 times and every 5 time steps all distances reported by scouts and recruits were recorded and combined. This was done to reflect the way foraging data is collected in real honeybee studies. We fit an exponential and minimum of an exponential distribution to both the distribution of foraging distances reported by the scout and recruit objects. Fitting was done by deriving the maximum likelihood estimate for each model fit on each data source through their analytical solutions: $\hat{\lambda} = \frac{1}/{\bar{x}}$, minimum of the exponential with a minimum foraging distance: $\hat{\lambda} = \frac{1}{\pi \bar{x^{2}}}$. As the exponential assumes distributions start from 0 the data was transformed to start from 0 by subtracting the minimum foraging distance from all foraging distances ($x = x - min(x)$) before fitting.

@@

We implemented this in a full model which describes the distance distribution of the environment has $n$ different resource types (See Supplementary Material). However, in the full model the number of parameters increases with $2n$.  Even if the number of resources is low, it turned out to be cumbersome to estimate the parameters and the model tends to overfit. To facilitate estimation of the parameter $p$ we therefore used a simplified model to estimate the fraction of scout trips.

@@

$$ f_s(x)=M_s^{-1} a_s b_s e^{-b_s a_s x}  \left[ 1-a_s x\right]_+ ,$$

where we used the shorthand ${x}_+=\max(x,0).$ The parameter $a_s^{-1}$ is the maximum reported distance by scouts, $a_s b_s$ is the intensity of resources found by scouts and $M_s=1-b_s^{-1} \left(1-e^{-b_s}\right)$ is factor that normalises the distribution.

@@

$$f_r(x)=M_r^{-1} 2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2}\left[1- a_r x\right]_+$$

where $M_r={1-\frac{erf\left(  \sqrt{\pi b_r}\right)}{2 \sqrt{b_r }}}$ is the normalisation factor, the parameter $a_r$ is the rate with which dances repeats depends on distance for recruit trips and $a_r^2 b_r$ is the intensity of high quality resources reported by recruited foragers.

@@

Goodness of fit is assessed using the Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package 'Matching' in R [@Sekhon2011].

@@

As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we performed a beta regression using the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010]. As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}, as outlined in the betareg package documentation. After analysis the data was back transformed to the original values for the plots in Fig 5.

>>>>

A circular environment is first created with radius $r = 2.5$. The number of flowers in the environment is generated as a random Poisson variable with rate equal to 5000 multiplied by the area of the environment. These are placed on polar coordinates with a uniformly selected angle, $\theta$, between 0 and 2$\pi$ and a radial value, $\rho$, between 0 and $r$, determined from the square root of the uniform position values multiplied by $r$. These polar coordinates are converted to Cartesian coordinates using the formulas $x = \rho\cos(\theta)$ and $y = \rho\sin(\theta)$. Each location is then assigned to an instance of a flower object along with a random quality of between 0 and 10. This quality is combined with the distance of the flower to the centrally located hive to form a measure of how profitable the flower is (see model, equation?).

@@

The simulation was run 100 times and every 5 time steps all distances reported by scouts and recruits were recorded and combined. This was done to reflect the way foraging data is collected in real honeybee studies. We fit an exponential and minimum of an exponential distribution to both the distribution of foraging distances reported by the scout and recruit objects. Fitting was done by deriving the maximum likelihood estimate for each model fit on each data source through their analytical solutions: $\hat{\lambda} = \frac{1}{\bar{x}}$, minimum of the exponential with a minimum foraging distance: $\hat{\lambda} = \frac{1}{\pi \bar{x^{2}}}$. As the exponential assumes distributions start from 0 the data was transformed to start from 0 by subtracting the minimum foraging distance from all foraging distances ($x = x - min(x)$) before fitting.

@@

We implemented this in a full model which describes the distance distribution of the environment has $n$ different resource types (See Supplementary Material). However, in the full model the number of parameters increases with $2n$.  Even if the number of resources is low, it turned out to be cumbersome to estimate the parameters and the model tends to overfit. To facilitate estimation of the parameter $p$ we therefore used a simplified model to estimate the fraction of scout trips, where $m$ represents the lowest duration recorded, as there exists a minimum waggle run duration.

@@

$$ f_s(x)=a_s \left[ M_s^{-1} b_s e^{-b_s a_s (x-m)}  \left[ 1-a_s x \right]_+ \right] $$

where we used the shorthand ${x}_+=\max(x,0)$ The parameter $a_s^{-1}$ is the maximum reported distance by scouts, $a_s b_s$ is the intensity of resources found by scouts and $M_s=(1-a_s m)-b_s^{-1} \left(1-e^{-b_s(1-a_s m)}\right)$ is factor that normalises the distribution.

@@

$$f_r(x)=M_r^{-1} 2 \pi a_r^2   b_r x e^{- \pi  b_r  (a_r x)^2} \left[1- a_r x\right]_+$$

where

$$M_r={(1-a_r m)e^{-\pi br(a_r m)^2} + \frac{erf(a_r \sqrt{\pi b_r m}) - erf(\sqrt{\pi b_r}))}{2 \sqrt{b_r}}}$$

is the normalisation factor, the parameter $a_r$ is the rate with which dances repeats depends on distance for recruit trips and $a_r^2 b_r$ is the intensity of high quality resources reported by recruited foragers.

@@

Goodness of fit is assessed using the two-sample Kolmorgorov-Smirnov (KS) test [@Goldstein2004] and implemented in R using the ks.boot function of the package 'Matching' in R [@Sekhon2011].

@@

Prior to conducing the PLS we removed any sites in which the models fit was significantly different to the waggle dance durations for that site. This resulted in one agri-rural site (ROT) and no urban sites being removed from the analysis.

As our estimated proportion of scouts is continuos on the interval $[0, 1]$ we used the R package plsRbeta [@Bertrand2013] to conduct the partial least squares analysis and performed a beta regression on the results using the R package betareg [@Cribari-Neto2010]. As the betareg package only works on the open interval $(0, 1)$ the data, $x$, was transformed using the following equation: $\frac{x(n-1)+0.5}{n}, as outlined in the betareg package documentation. After analysis the data was back transformed to the original values for the plots in Fig 5.
